\section{Package overview}
The current contents of HmiGraphics and HmiAnimation is as follows:

\begin{itemize}
\item hmi.animation:
Deals with virtual objects, in particular animation of such objects. It defines a simple hierarchy of
virtual objects, by means of VJoints, primarily used for skeleton based animation of human avatars.
a virtual environment. Of course it is intended to be used in combination with the hmi.graphics packages,
but in principle it can be build and used independently.
\item hmi.graphics.collada:
Deals with reading (and to some extent writing) graphics files according to the Collada standard.
Apart from reading, it includes a sub package hmi.graphics.collada.scenegraph, for translating Collada descriptions into
hmi.graphics.scenegraph descriptions.
\item hmi.graphics.geometry:
Utility class, dealing with some purely geometric algorithms.
\item hmi.graphics.opengl:
Deals with actual rendering OpenGL based 3D graphics.
%It defines counterparts of some classes of the
%scenegraph package. For instance, we have the GLBasicMesh class which is sort of the hmi.graphics.opengl counterpart
%for the hmi.graphics.scenegraph GMesh class.
Although hmi.graphics.opengl is OpenGL based, it is still independent
of the  particular Java OpenGL binding or implementation. Currently there are two such OpenGL bindings: Jogl and LWJGL.
Each of these has its own package (hmi.graphics.jogl, hmi.graphics.lwjgl), and one of them must be used in combination with hmi.graphics.opengl.
hmi.graphics.opengl itself has a few sub packages:
\begin{itemize}
\item hmi.graphics.opengl.state: Contains classes for dealing with the OpenGL state management.
\item hmi.graphics.opengl.geometry: Defines a few utility classes for rendering simple geometry, like spheres, boxes, and lines.
\item hmi.graphics.opengl.scenegraph: Defines the translation from hmi.graphics.scengraph structures to hmi.graphics.opengl structures.
\end{itemize}

\item hmi.graphics.jogl and hmi.graphics.lwjgl are two packages that provide
actual OpenGL bindings, relying on native code in the form of windows dll files or Linux so files.
Both packages implement the GLRenderContext interface from hmi.graphics.opengl, and define
a simple Renderer.
Most of the JOGLContext and LWJGLContext code is generated, by means of a utility package hmi.graphics.gen.


\item hmi.graphics.scenegraph:
Defines scenegraph like structures for representing graphics data in a format independent of the external file format,
but also independent of the actual render technology being used.
\item hmi.graphics.util:
The usual package of our favorite ``utilities''.

\end{itemize}

Currently, there are a few more ``deprecated'' and/or temporary packages:
\begin{itemize}
\item hmi.graphics.colladatest:
Just for testing packages like collada, opengl, etcetera.
\item hmi.graphics.gen: used for generating files within  jogl and lwjgl implementations.
\item hmi.graphics.render
An attempt to create a platform neutral 3D renderer.
\end{itemize} 